Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-11 paper/paper.tex

\begin{abstract}
Abstracting abstract machines has recently been proposed as a
lightweight approach to designing sound and computable program
analyses.  The approach involves deriving abstract interpreters from
existing machine semantics and has been applied to a variety of
languages with features widely considered difficult to analyze.
Although analyzers are straightforward to build under this approach,
the resulting analyzer is prohibitively inefficient.

@@

The \emph{abstracting abstract machines}
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides a straightforward way of
transforming a programming language semantics in the form of an
abstract machine, into a family of abstract interpreters parameterized
over policies for regulating the analytic precision of the control,
environment, store, and base value domains.  By taking a
machine-oriented view of computation, it becomes possible to design,
verify, and implement program analyzers for realistic language
features typically considered difficult to model.  The approach was
originally applied to features such as higher-order functions,
stack-inspection, exceptions, laziness, first-class continuations, and
garbage collection.  It has since been used to verify actor-
\cite{local:DOsualdo:12A} and
thread-based~\cite{dvanhorn:Might2011Family} parallelism and
behavioral contracts~\cite{dvanhorn:TobinHochstadt2012Higherorder}; it
has been used to model Coq~\cite{local:harvard},
Dalvik~\cite{local:dalvik}, Erlang~\cite{local:DOsualdo:12B},
JavaScript~\cite{local:DBLP:journals/corr/abs-1109-4467}, and
Racket~\cite{dvanhorn:TobinHochstadt2012Higherorder}.

@@

We start, in section~\ref{sec:aam}, by developing an abstract
interpreter according to the abstracting abstract machines approach
and observe the resulting implementation cannot analyze the example in
under 30 minutes.  In section~\ref{sec:baseline}, we perform a further
abstraction by widening the store.  The resulting analyzer sacrifices
precision for speed and is able to analyze the example in about 1
minute.  Without this abstraction step the analysis is exponential due
to per-state store variance.  This step is described by Van Horn and
Might and is necessary to make even small examples feasible.  We
therefore take the widened interpreter as the baseline for our
evaluation.

@@

>>>>

\begin{abstract}
Abstracting abstract machines has recently been proposed as a
lightweight approach to designing sound and computable program
analyses.  The approach involves deriving abstract interpreters from
existing machine semantics and has been applied to a variety of
languages with features widely considered difficult to analyze.
Although analyzers are straightforward to build under this approach,
the result is prohibitively inefficient.

@@

The \emph{abstracting abstract machines} (AAM)
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides a straightforward way of
transforming a programming language semantics in the form of an
abstract machine, into a family of abstract interpreters parameterized
over policies for regulating the analytic precision of the control,
environment, store, and base value domains.  By taking a
machine-oriented view of computation, it becomes possible to design,
verify, and implement program analyzers for realistic language
features typically considered difficult to model.  The approach was
originally applied to features such as higher-order functions,
stack-inspection, exceptions, laziness, first-class continuations, and
garbage collection.  It has since been used to verify actor-
\cite{local:DOsualdo:12A} and
thread-based~\cite{dvanhorn:Might2011Family} parallelism and
behavioral contracts~\cite{dvanhorn:TobinHochstadt2012Higherorder}; it
has been used to model Coq~\cite{local:harvard},
Dalvik~\cite{local:dalvik}, Erlang~\cite{local:DOsualdo:12B},
JavaScript~\cite{local:DBLP:journals/corr/abs-1109-4467}, and
Racket~\cite{dvanhorn:TobinHochstadt2012Higherorder}.

@@

We start, in section~\ref{sec:aam}, by developing an abstract
interpreter according to the abstracting abstract machines approach
and observe the resulting implementation cannot analyze the example in
under 30 minutes.  In section~\ref{sec:baseline}, we perform a further
abstraction by widening the store.  The resulting analyzer sacrifices
precision for speed and is able to analyze the example in about 1
minute.  Without this abstraction step the analysis is exponential due
to per-state store variance.  This step is described by Van Horn and
Might~\cite[\S 3.5--6]{dvanhorn:VanHorn2012Systematic} and is
necessary to make even small examples feasible.  We therefore take the
widened interpreter as the baseline for our evaluation.

@@

\paragraph{Acknowledgments}

We thank Suresh Jagannathan for providing source code to the
polymorphic splitting
analyzer~\cite{dvanhorn:wright-jagannathan-toplas98} and Ilya Sergey
for the introspective pushdown
analyzer~\cite{dvanhorn:Earl2012Introspective}.
